---
title: "data_manip"
author: "Maggie Li (ml4424)"
date: "9/22/2020"
output: html_document
---

## load library(s)
```{r}
library(tidyverse)
```

## Read in datasets
```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data/data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/data_import_examples/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## select function
```{r}
# only keep certain cols
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)

# piping looks better tbh
litters_data %>% select(group, litter_number, gd0_weight, pups_born_alive)

# specify range of cols to keep
litters_data %>% select(group:gd_of_birth)

# to remove
litters_data %>% select(-pups_survive)

# can rename cols WHILE selecting (OMG!)
litters_data %>% select(GROUP = group, LiTtEr_NuMbEr = litter_number)

# using rename only w/o selecting
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)

# helpers for select(); kind of like filtering cols that have specific starts, ends and contains
select(litters_data, starts_with("gd"))

# everything() arg in select() function pretty much just moves columns around, the first ones being the other args, and keeps the rest of the cols in data table
litters_data
select(litters_data, litter_number, pups_survive, everything())

# relocate() is a function that does the same thing as above
relocate(litters_data, litter_number, pups_survive) 

```

## Filter function
Pretty straightforward, you've done it before but check notes for add examples

## Cmd --> Shift --> M does piping !

## Mutate function
This creates new columns based on existing ones

```{r}
litters_data %>% mutate(
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group) #converts entries to all lower case
)

# create variable that subtracts 7 from PD pivot
pups_data %>% mutate(
  sub_7 = pd_pivot - 7,
  sum_pd = pd_ears + pd_eyes + pd_pivot + pd_walk
)
```
A few things in this example are worth noting:

-Your new variables can be functions of old variables
-New variables appear at the end of the dataset in the order that they are created
-You can overwrite old variables
-You can create a new variable and immediately refer to (or change) it


## final learning assessment
```{r}
pups_data_final <- read_csv("./data/data_import_examples/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  filter(sex == 1) %>% 
  select(-pd_ears) %>% 
  mutate(pd_7 = ifelse(pd_pivot >= 7, 1, 0)) #can also just be mutate(pd_7 = pd_pivot > 7), don't need ifelse

```

